### 2024-03-21 paper/assets/features_table.tex
Initial paper draft

\begin{table}
\centering
\caption{Features used in modeling and analysis}
\label{tab:features}
\begin{tabular}{llll}
\toprule
 & feature & entity & kind \\
\midrule
1 & disease\_\_clinical\_\_phase\_max\_\_reached & disease & temporal \\
2 & target\_\_clinical\_\_phase\_max\_\_reached & target & temporal \\
3 & target\_\_genetic\_constraint & target & static \\
4 & target\_\_mouse\_ko\_score & target & static \\
5 & target\_\_tissue\_distribution & target & static \\
6 & target\_\_tissue\_specificity & target & static \\
7 & target\_disease\_\_affected\_pathway\_\_cancer\_biomarkers & target\_disease & temporal \\
8 & target\_disease\_\_affected\_pathway\_\_crispr & target\_disease & temporal \\
9 & target\_disease\_\_affected\_pathway\_\_crispr\_screen & target\_disease & temporal \\
10 & target\_disease\_\_affected\_pathway\_\_progeny & target\_disease & temporal \\
11 & target\_disease\_\_affected\_pathway\_\_reactome & target\_disease & temporal \\
12 & target\_disease\_\_affected\_pathway\_\_slapenrich & target\_disease & temporal \\
13 & target\_disease\_\_affected\_pathway\_\_sysbio & target\_disease & temporal \\
14 & target\_disease\_\_animal\_model\_\_impc & target\_disease & temporal \\
15 & target\_disease\_\_genetic\_association\_\_clingen & target\_disease & static \\
16 & target\_disease\_\_genetic\_association\_\_curated & target\_disease & static \\
17 & target\_disease\_\_genetic\_association\_\_eva & target\_disease & static \\
18 & target\_disease\_\_genetic\_association\_\_gene2phenotype & target\_disease & static \\
19 & target\_disease\_\_genetic\_association\_\_gene\_burden & target\_disease & static \\
20 & target\_disease\_\_genetic\_association\_\_genomics\_england & target\_disease & static \\
21 & target\_disease\_\_genetic\_association\_\_omim & target\_disease & static \\
22 & target\_disease\_\_genetic\_association\_\_orphanet & target\_disease & static \\
23 & target\_disease\_\_genetic\_association\_\_ot\_genetics\_portal & target\_disease & static \\
24 & target\_disease\_\_genetic\_association\_\_uniprot\_literature & target\_disease & static \\
25 & target\_disease\_\_genetic\_association\_\_uniprot\_variants & target\_disease & static \\
26 & target\_disease\_\_known\_drug\_\_chembl & target\_disease & temporal \\
27 & target\_disease\_\_literature\_\_europepmc & target\_disease & temporal \\
28 & target\_disease\_\_outcome\_\_advanced & target\_disease & temporal \\
29 & target\_disease\_\_rna\_expression\_\_expression\_atlas & target\_disease & temporal \\
30 & target\_disease\_\_somatic\_mutation\_\_cancer\_gene\_census & target\_disease & temporal \\
31 & target\_disease\_\_somatic\_mutation\_\_eva\_somatic & target\_disease & temporal \\
32 & target\_disease\_\_somatic\_mutation\_\_intogen & target\_disease & temporal \\
33 & target\_disease\_\_time\_\_transition & target\_disease & temporal \\
\bottomrule
\end{tabular}
\end{table}